Fail Faster

### 2025-03-18 paper/paper.tex
progress
chart

  % in strict functionallanguages like OCaml and Scal
  To address this question, we provide two key insights for designing generator libraries, which eliminates unnecessary overhead while preserving expressiveness.
  First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used for
optimizing functional DSLs~\cn.
We apply this technique to generators, completely eliminating the overhead of
many common generator abstractions. For example, in unstaged generator
libraries, each call to monadic bind usually spends time constructing and
immediately executing a closure at run time. \jwc{Todo: joe --- cite some evidence for this}
Meanwhile, a staged approach can resolve the bind at compile time, enabling the
compiler to inline and eliminate these closures.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison. \iflater\hg{More words here for
parallelism?}\fi Together, these insights provide a recipe for improving existing generator libraries.

@@

\subsection{Implementing and Testing Generators}
\label{subsection:pbt-for-pbt}
To establish the equivalence of generators up to program equality, 
we employ a PBT harness to compare their outputs over 1,000 random seeds. 
If a discrepancy arises, we report the unequal trees; if no counterexample is found,
we conclude that the generators exhibit equivalent behavior.
\subsection{Benchmarking speed \& resource usage}

>>>>

  % in strict functionallanguages like OCaml and Scal
  To address this question, we outline two key principles for designing generator libraries that balance efficiency with flexibility.
 First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used for
optimizing functional DSLs~\cn.
We apply this technique to generators, completely eliminating the overhead of
many common generator abstractions. For example, in unstaged generator
libraries, each call to monadic bind usually spends time constructing and
immediately executing a closure at run time. \jwc{Todo: joe --- cite some evidence for this}
Meanwhile, a staged approach can resolve the bind at compile time, enabling the
compiler to inline and eliminate these closures.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison. \iflater\hg{More words here for
parallelism?}\fi Together, these insights provide a recipe for improving existing generator libraries.

@@

\subsection{Implementing and Testing Generators}
\begin{figure}[h]
  \includegraphics[scale=0.34]{allocs.png}\includegraphics[scale=0.334]{time.png}
\end{figure}
\label{subsection:pbt-for-pbt}
To establish the equivalence of generators up to program equality, 
we employ a PBT harness to compare their outputs over 1,000 random seeds. 
If a discrepancy arises, we report the unequal trees; if no counterexample is found,
we conclude that the generators exhibit equivalent behavior.
\subsection{Benchmarking speed \& resource usage}
